Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Saturn-Sized Planet in IC~2602

### 2020-07-10 paper/ms.tex
update manuscript

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, photometry from El Sauce
  and ASTEP400,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in the open cluster IC 2602.  The
  star and validated planet are therefore $35^{+11}_{-5}$ million
  years old.  The planet is warm ($P = 8.3\,{\rm d}$) and slightly
  smaller than Jupiter.  Combining the transit photometry with a prior on the
  stellar parameters derived from the cluster color-magnitude diagram,
  we find that the transits are grazing with $b=0.93\pm0.01$ and
  $R_{\rm p} = 0.79 \pm 0.07 \,R_{\rm Jup}$.  From radial velocity
  monitoring, we limit the minimum mass to less than 1.20 $M_{\rm
  Jup}$ (3$\sigma$).  Grazing transits are a cause for concern, as
  they are usually indicative of astrophysical false positive
  scenarios.  Our follow-up data show that such scenarios are highly
  unlikely.  Our combined multi-color photometry, high-resolution
  imaging, and radial velocities rule out hierarchical eclipsing
  binary scenarios.  Background eclipsing binary scenarios, though
  limited by speckle imaging, formally remain a 0.2\% possibility.
  TOI 837.01 is therefore a validated adolescent exoplanet.  Further
  observations of its stellar obliquity, atmosphere, and mass should
  confirm its planetary nature beyond doubt, and could improve our
  understanding of how the physical and orbital properties of
  exoplanets change in time.
\end{abstract}

@@

%TODO: you must check custom lightkurve apertures, too

% %FIXME: this paragraph doesn't really belong here?
% We examined the CDIPS FFI light curves of the target, which are
% available on MAST \citep{bouma_cluster_2019}. Three light curves are
% available, based on photometric apertures with a radius of 1, 1.5, or
% 2.5 pixels.  In the raw difference-image light-curves, as well as the
% PCA-detrended light-curves, dips of depth $\approx$0.35\% are visible,
% and their properties do not significantly vary with aperture size.
% Within $\approx20$'', the dips are consistent with originating from
% the target star.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene of \tn.}
    {\it Top:} Mean TESS image of \tn\ over Sector~10, with a
    logarithmic grayscale. The yellow star is the position of \tn.
    White circles are resolved Gaia sources with $T<16$ (brighter stars
    are larger). The black \texttt{X} and white \texttt{/} hatches
    show the apertures used to measure the background and target star
    flux, respectively. Dashed lines of constant declination and right
    ascension are shown.
    Two stars of interest are ``Star A'' and ``Star B'', which were eventually excluded as being possible sources of the transits.
    %{\it Bottom:} Digitized Sky Survey $R$-band
   % image of the same field, with a linear grayscale. 
		\label{fig:scene}
	}
\end{figure}

@@

The Gaia imaging, reduced to its point-source catalog, provides
the initial context for analyzing the TESS data.  Stars brighter than
$T=16$, as queried from the Gaia DR2 source catalog, are shown with
white circles in Figure~\ref{fig:scene}, overlaid on 
the TESS image.  Given its galactic latitude of $b=-6^\circ$, it is
not surprising that the field of \tn\ is crowded.  The resolved stars
that were of immediate concern for our false positive analysis were as
follows.
\begin{itemize}
  %
  \item \tn\ $\equiv$ TIC 460205581 ($T=9.9$). The target star.
  %
  \item Star A $\equiv$ TIC 847769574 ($T=14.6$), $2.3''$ west. The
    proper motions and parallax of this star imply it is comoving with
    \tn and is separated from it by $6.6\pm 0.1\,{\rm pc}$.
    It is therefore likely an \cn\ member, but unlikely to be a bound
    binary compnaion.
  %
  \item Star B $\equiv$ TIC 460205587 ($T=13.1$), $5.4''$ north.  The
    Gaia parallax implies this is a giant background star.
\end{itemize}
An additional source, TIC 847769581, is $4.9''$ from the target, but
too faint ($T=18.8$) to be the source of the observed transit signal.

The Gaia DR2 data for Star A seems poorly behaved.  While Star A has
$G=15.1$, and $Bp=14.9$, no $Rp$ magnitude is reported.
Correspondingly, no RUWE\footnote{ See the Gaia DPAC technical note
GAIA-C3-TN-LU-LL-124-01,
\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412},
\texttt{2020-07-08}. } value is available.  We suspect that the photometric
failure to produce an $Rp$ magnitude, as well as its poor astrometric
fit, are likely due to blending with \tn.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.47\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Ground-based follow-up photometry.} The data were acquired
		using the 0.36m at El Sauce and the 0.40m ASTEP400 at Dome C. Black points are
		binned to 10-minute intervals.  The gray line is the best-fit
		model to the combined TESS and ground-based data. Models show
		$2\sigma$ lower limits on the transit depths in the Cousins-R and
		Johnson-B bandpasses used to rule out specific false positive
		scenarios (see Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

The easiest path towards definitively confirming whether \tn\ is a
planet will be a Rossiter-McLaughlin (RM) measurement.  Detection of
the RM signal would rule out BEB and HEB scenarios, as it would imply
that the eclipsing object is bound to the target star. Combined with
our non-detection of the planet's mass from radial velocity
monitoring, this would confirm that \pn\ is a planet.

@@

%\clearpage
\acknowledgements
\tn\ was included on the TESS ``short-cadence'' target list in part
thanks to the Guest Investigator program of G.\ Sacco (G011265).
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
The Digitized Sky Survey was produced at the Space Telescope Science
Institute under U.S. Government grant NAG W-2166.
Figure~\ref{fig:scene} is based on photographic data obtained using
the Oschin Schmidt Telescope on Palomar Mountain.
%
This research made use of the Exoplanet Follow-up Observation
Program website, which is operated by the California Institute of
Technology, under contract with the National Aeronautics and Space
Administration under the Exoplanet Exploration Program.
%
This research has made use of the SVO Filter Profile Service
(\url{http://svo2.cab.inta-csic.es/theory/fps/}) supported from the Spanish
MINECO through grant AYA2017-84089.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

>>>>

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, photometry from El Sauce
  and ASTEP400, and spectroscopy from CHIRON, FEROS, and Veloce.  We
  find that TOI$\,$837 is a $T=9.9$ G0 dwarf in the open cluster IC
  2602.  The star and planet are therefore $35^{+11}_{-5}$
  million years old.  The planet is warm ($P = 8.3\,{\rm d}$) and
  slightly smaller than Jupiter.  Combining the transit photometry
  with a prior on the stellar parameters derived from the cluster
  color-magnitude diagram, we find that the transits are grazing with
  $b=0.93\pm0.01$ and $R_{\rm p} = 0.79 \pm 0.07 \,R_{\rm Jup}$.  From
  radial velocity monitoring, we limit the minimum mass to less than
  1.20 $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are a cause for
  concern, as they are usually indicative of astrophysical false
  positive scenarios.  Our follow-up data show that such scenarios are
  highly unlikely.  Our combined multi-color photometry,
  high-resolution imaging, and radial velocities rule out hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios,
  though limited by speckle imaging, formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent
  exoplanet.  Further observations of its stellar obliquity,
  atmosphere, and mass should confirm its planetary nature beyond
  doubt, and could improve our understanding of how the physical and
  orbital properties of exoplanets change in time.
\end{abstract}

@@

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene of \tn.}
    Mean TESS image of \tn\ over Sector~10, with a logarithmic
    grayscale. The yellow star is the position of \tn.  White circles
    are resolved Gaia sources with $T<16$ (brighter stars are larger).
    The black \texttt{X} and white \texttt{/} hatches show the
    apertures used to measure the background and target star flux,
    respectively. Dashed lines of constant declination and right
    ascension are shown.  Two stars of interest are ``Star A'' and
    ``Star B'', which were eventually excluded as being possible
    sources of the transits.
		\label{fig:scene}
	}
\end{figure}

@@

The Gaia imaging, reduced to its point-source catalog, provides the
initial context for analyzing the TESS data.  Stars brighter than
$T=16$, as queried from the Gaia DR2 source catalog, are shown with
white circles in Figure~\ref{fig:scene}, overlaid on the TESS image.
Given its galactic latitude of $b=-6^\circ$, it is not surprising that
the field of \tn\ is crowded.  The resolved stars that were of
immediate concern for our false positive analysis were as follows.
\begin{itemize}
  %
  \item \tn\ $\equiv$ TIC 460205581 ($T=9.9$). The target star.
  %
  \item Star A $\equiv$ TIC 847769574 ($T=14.6$), $2.3''$ west. The
    proper motions and parallax of this star imply it is comoving with
    \tn\ and that the two stars are separated by $6.6\pm 0.1\,{\rm pc}$.
    Star A is therefore likely an \cn\ member, but unlikely to be a bound
    binary compnaion.
  %
  \item Star B $\equiv$ TIC 460205587 ($T=13.1$), $5.4''$ north.  The
    Gaia parallax implies this is a giant background star.
\end{itemize}
An additional source, TIC 847769581, is $4.9''$ from the target, but
too faint ($T=18.8$) to be the source of the observed transit signal.

The Gaia DR2 data for Star A seems poorly behaved.  While Star A has
$G=15.1$, and $Bp=14.9$, no $Rp$ magnitude is reported.
Correspondingly, no RUWE\footnote{ See the Gaia DPAC technical note
GAIA-C3-TN-LU-LL-124-01,
\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412},
\texttt{2020-07-08}. } value is available.  We suspect that the photometric
failure to produce an $Rp$ magnitude as well as its poor astrometric
fit are due to blending with \tn.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.49\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Ground-based follow-up photometry.} The data were acquired
		using the 0.36m at El Sauce and the 0.40m ASTEP400 at Dome C. Black points are
		binned to 10-minute intervals.  The gray line is the best-fit
		model to the combined TESS and ground-based data. Models show
		$2\sigma$ lower limits on the transit depths in the Cousins-R and
		Johnson-B bandpasses used to rule out specific false positive
		scenarios (see Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

The easiest path towards definitively confirming whether \tn\ is a
planet will be a Rossiter-McLaughlin (RM) measurement.  Detection of
an RM signal consistent with the photometric transit would rule out
BEB and HEB scenarios, as it would imply that the eclipsing object is
bound to the target star. Combined with our non-detection of the
planet's mass from radial velocity monitoring, this would confirm that
\pn\ is a planet.

@@

%\clearpage
\acknowledgements
\tn\ was included on the TESS ``short-cadence'' target list in part
thanks to the Guest Investigator program of G.\ Sacco (G011265).
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%
This research made use of the Exoplanet Follow-up Observation
Program website, which is operated by the California Institute of
Technology, under contract with the National Aeronautics and Space
Administration under the Exoplanet Exploration Program.
%
This research has made use of the SVO Filter Profile Service
(\url{http://svo2.cab.inta-csic.es/theory/fps/}) supported from the Spanish
MINECO through grant AYA2017-84089.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%
